### 2016-09-14 cross_compilation.tex
first round of corrections

\label{sec:cross_compilation}
An important case of a computational expensive task is to compile software
packages and large libraries. Given the computing capabilities of the
\ac{Raspi}, such tasks can be challenging if not prohibitive in terms of
\ac{CPU}, memory or space usage and / or compilation time. Thus, in
this Section we present a procedure of how to cross-compile C++
source code from the testbed administrator \ac{PC} for the \ac{ARM}
architecture of the \ac{Raspi}s. By doing this, we take advantage
of the (typically) much higher computing power of the testbed
administrator \ac{PC} in order to save time and computational resources.
Hence, we give an example of compiling a simple C++ program and
copying the generated binaries with \ac{SSH} to run locally
on a \ac{Raspi}.

@@

\subsection{Toolchain Setup}
To compile in a given architecture that is aimed for a different one,
the testbed administrator needs to install a toolchain in its \ac{PC}.
The toolchain is mandatory due to the different processor architectures
where the source can be compiled from.
Given that compiling a toolchaing can be an ardous task, we get the toolchain
recommended for the \ac{ARM} architecture of the \ac{Raspi}s.
This toolchain is available from \cite{2016rpitoolchain} and
it already contains the binaries for different compilers based on
\texttt{gcc 4.9}. In what follows, we extract the binaries adjusting
them to our coding style and compiling convention. For this, we use the
\texttt{\$\{TOOLCHAIN\}} directory as the working directory. However,
the testbed administrator may choose some other working directory of
its preference if desired. First, we create the toolchain directory:

@@

Later, we download the \ac{Raspi} toolchain for 64-bit Linux. However,
instead of working with the binaries from \cite{2016rpitoolchain},
we use a zipped version with the original binaries and renamed symbolic
links to the original files available in
\cite{2016steinwurfrpitoolchain}. Finally, we unzip the downloaded file.
This is made as follows:

@@

To use Kodo for research, it is required to obtain a research free license.
To do this, only a request form needs to be filled
in~\cite{steinwurflicenselink} and wait for grant access. Once
the access for Kodo has been granted, the source code can be pulled from
its Git repositories to be compiled. Thus, assuming that testbed
administrator already has access, we clone the \texttt{kodo-cpp}
repository locally in \texttt{\${CODEDIR}} and change directory into
the repository by doing:

@@

\begin{lstlisting}[]
$ ssh pi@<RASP_IP>
$ ./kodocpp_tests
...
[----------] Global test environment tear-down
[==========] 17 tests from 15 test cases ran. (XXXX ms total)
[  PASSED  ] 17 tests.
$ ./encode_decode_simple
Data decoded correctly
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

>>>>

\label{sec:cross_compilation}
An important case of a computational expensive task is to compile software
packages and large libraries. Given the computing capabilities of the
\ac{Raspi}, such tasks can be challenging if not prohibitive in terms of
\ac{CPU}, memory or space usage and / or compilation time. In
this section we present a procedure of how to cross-compile C++
source code from the testbed administrator \ac{PC} for the \ac{ARM}
architecture of the \ac{Raspi}s. By doing this, we take advantage
of the (typically) much higher computing power of the testbed
administrator \ac{PC} in order to save time and computational resources.
Hence, we give an example of compiling a simple C++ program and
copying the generated binaries with \ac{SSH} to run locally
on a \ac{Raspi}.

@@

\subsection{Toolchain Setup}
To compile in a given architecture that is aimed for a different one,
the testbed administrator needs to install a toolchain in its \ac{PC}.
The toolchain is mandatory due to the different processor architectures
where the source can be compiled from.
Given that compiling a toolchaing can be an ardous task, we get the toolchain
recommended for the \ac{ARM} architecture of the \ac{Raspi}s.
This toolchain is available from \cite{2016rpitoolchain} and
it already contains the binaries for different compilers based on
\texttt{gcc 4.9}. In what follows, we extract the binaries adjusting
them to our coding style and compiling convention. For this, we use the
\texttt{\$\{TOOLCHAIN\}} directory as the working directory.
The testbed administrator may choose some other working directory of
its preference if desired. First, we create the toolchain directory:

@@

Later, we download the \ac{Raspi} toolchain for 64-bit Linux.
Instead of working with the binaries from \cite{2016rpitoolchain},
we use a zipped version with the original binaries and renamed symbolic
links to the original files available in
\cite{2016steinwurfrpitoolchain}. Finally, we unzip the downloaded file.
This is made as follows:

@@

To use Kodo for research, it is required to obtain a research free license.
To do this, a request form needs to be filled
in~\cite{steinwurflicenselink} and wait for it to be processed by the Kodo
developers. Once
the access for Kodo has been granted, the source code can be pulled from
its Git repositories to be compiled. Assuming that the testbed
administrator already has access, we clone the \texttt{kodo-cpp}
repository locally in \texttt{\${CODEDIR}} and change directory into
the repository by doing:

@@

\begin{lstlisting}[]
$ ssh pi@<RASP_IP>
$ ./kodocpp_tests
...
[  PASSED  ] X tests.
$ ./encode_decode_simple
Data decoded correctly
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

### 2016-09-14 image_setup.tex
first round of corrections

To get started, we first need to install an \ac{OS} that works properly
on all the \ac{Raspi} models. We will download and setup the image in
the testbed administrator \ac{PC} using a Debian-based distribution. We
use the popular Debian-based Raspbian Linux~\cite{raspbian} given that is
the recommended and default \ac{OS} for the \ac{Raspi}. Raspbian is made
available in two bundles: Raspbian and Raspbian Lite. The difference
between the two, is that Raspbian contains a pre-installed desktop environment
for user interaction and Raspbian Lite by default only permits to interact
through a command shell. Given that the \ac{Raspi}s in our testbed are not
connected to monitors, we decide to work with Raspbian Lite. Still, if later
desired, a desktop environment can be installed using the package manager
as it will be shown in the configuration section.

@@

After Raspbian Lite has been unpacked, there should be an \texttt{.img}
file in the working directory \texttt{\$\{WORKDIR\}}. Here, \texttt{fdisk} can
be used to display the content of the image. We parse the arguments \texttt{-u sectors}
to display the sizes in sectors and \texttt{-l} to display the partitions
within the image. The \texttt{fdisk} command should output to the terminal
something similar to:

@@

\subsection{Block Devices Mounting}
For browsing and altering the files in the image, we mount the block
device partitions into a particular path of our \texttt{\$\{WORKDIR\}} in order
to customize them. We proceed first to mount the block device partition that
contains the root filesystem and later the boot partition. This is done by
creating an empty directory that is used as a mountpoint. We name it
\texttt{root} and create it in the working directory before mounting the
root filesystem onto the mountpoint. Thus, we mount the root filesystem
as follows:

@@

\subsection{Image OS Files and Configuration Scripts Setup}
In general, the \ac{Raspi}s should be setup as similar as possible. However,
some particularities exist to differentiate the devices in principle. Also,
scripts containing further configurations for the \ac{Raspi}s are desirable
to be distributed as part of the common image. Therefore, here we present
the steps to setup basic properties of the \ac{Raspi}s and distributing
configuration scripts to each of them through the image. For this, we first
indicate how to obtain and put our configuration scripts in the image. Later,
we describe the tasks performed by these configuration scripts. Finally, we
indicate how and in which order the scripts are executed to configure all the
devices. Nevertheless, any testbed administrator might modify or include other
tasks besides the presented in this procedure according to his / her needs
as we will show.

@@

% MAC and Hostname file
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/nodes.csv}@>
<@\textcolor{gray}{
--------------------------------------------------------------------------------}
\Reactivatenumber @>
# Ethernet MAC    Hostname
b8:27:eb:5b:da:20 rasp00
b8:27:eb:7b:c3:91 rasp01
b8:27:eb:54:9c:64 rasp02
b8:27:eb:95:bd:11 rasp03
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

\subsubsection{Updating Default Configuration Files and Scripts}
Besides the single script with its configuration file introduced up
to this point of our procedure, it is possible that the testbed
administrator may require to add other scripts to configure
his / her \ac{Raspi}s. Hence, we want
to ensure that all the \ac{Raspi} configuration scripts of any testbed
administrators are obtained in a simple way. We automate this task by
including the \texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/update\_rasp\_config}
script in our procedure. The purpose of this script is to make all the
\ac{Raspi}s fetch all the configuration scripts located with the image
during a testbed start up.

In our case as the testbed administrator for presenting the procedure, we want
to fetch all our configuration scripts in
\texttt{\$\{URL\%/*\}/raspi\_config.zip}. Thus, the update script
\textit{automatically} downloads all the required configuration files in
\texttt{rasp\_config.zip} file from a remote location. This is the same that
we manually did earlier to get our files but this will be made in an
automated way after booting up the system. This script content is:

@@

The update script: (3-4) checks for the \ac{URL} and \texttt{.zip} file that
should be downloaded, (7-10) downloads the configuration files to \texttt{/tmp}
folder in the corresponding \ac{Raspi}. It also prints a warning in case of
errors and (13) unzips the files to \texttt{/home/pi/}, the \ac{Raspi} home
directory. Existing files and directories are simply overwritten.
%while ignoring existing files and / or directories there.

For the above scripts to work in the \ac{Raspi}s, it is required that the
\ac{Raspi}s \ac{MAC} addresses are found in \texttt{nodes.csv}. Also, it
should be noted that for other testbed administrators besides ourselves,
the \ac{URL} for file fetching and the configuration
scripts themselves can be modified to fit their requirements. Thus, if
required for a testbed administrator, the \texttt{rasp\_config.zip}
will need to be edited to include all the required configuration files
and scripts. Also, it might be needed to edit the \ac{URL} in the script
\texttt{update\_rasp\_config} to store and fetch from a different location.
Nevertheless, both the \ac{URL} and configuration files presented here can be
used as a starting boilerplate if desired.

\subsubsection{Configuration Scripts Execution Order}
To actually make the \ac{Raspi}s change hostnames and any other considered
configurations, we have to make each \ac{Raspi} call the above scripts when
it starts up. After finishing the setup process, all the unzipped files
presented in Section~\ref{sec:configuration_files_download} should be
locally available at each \ac{Raspi} after getting the root filesystem.
Therefore, we first need to run the update script before running
any other configuration scripts. To do this after boot up, we include
a call for the update script in the image \texttt{\$\{ROOTDIR\}/etc/rc.local}
file before \texttt{exit 0} using a file editor. To do this, it is required
to edit this file with root permissions. The resulting file should look like the following:

@@

Provided that qemu-arm is enabled, we should now be able to \texttt{chroot}
into our Raspbian lite image. However, there are a few commands to be
performed before actually changing root into the root partion of the image.
First, to get internet access from within the Raspbian lite image it is needed
to copy the testbed administrator Linux distribution \texttt{resolv.conf} file
into the image root filesystem. To do this, it is required to run the
following:

@@

\texttt{vim} is the improved \texttt{vi} editor for Linux, \texttt{git}
for managing Git repositories and \texttt{screen} for better handling of
long-runnning processes. When writing the image to a memory card, all the
changes that have been made to the image so far, will exist in all
\ac{Raspi}s after fetching it.

>>>>

To get started, we first need to install an \ac{OS} that works properly
on all the \ac{Raspi} models. We will download and setup the image in
the testbed administrator \ac{PC} using a Debian-based distribution. We
use the popular Debian-based Raspbian Linux~\cite{raspbian} given that is
the recommended and default \ac{OS} for the \ac{Raspi}. Raspbian is made
available in two bundles: Raspbian and Raspbian Lite. The difference
between the two, is that Raspbian contains a pre-installed desktop environment
for user interaction and Raspbian Lite by default only permits to interact
through a command shell. Given that the \ac{Raspi}s in our testbed are not
connected to monitors, we decide to work with Raspbian Lite.
A desktop environment can be installed using the package manager
as it will be shown in the configuration section.

@@

After Raspbian Lite has been unpacked, there should be an \texttt{.img}
file in the working directory \texttt{\$\{WORKDIR\}}. \texttt{fdisk} can
be used to display the content of the image. We parse the arguments \texttt{-u sectors}
to display the sizes in sectors and \texttt{-l} to display the partitions
within the image. The \texttt{fdisk} command should output to the terminal
something similar to:

@@

\subsection{Block Devices Mounting}
For browsing and altering the files in the image, we mount the block
device partitions into a particular path of our \texttt{\$\{WORKDIR\}} in order
to customize them. We mount the block device partition that
contains the root filesystem and later the boot partition. This is done by
creating an empty directory that is used as a mountpoint. We name it
\texttt{root} and create it in the working directory before mounting the
root filesystem onto the mountpoint. We mount the root filesystem
as follows:

@@

\subsection{Image OS Files and Configuration Scripts Setup}
In general, the \ac{Raspi}s should be setup as similar as possible. However,
some particularities exist to differentiate the devices in principle. Also,
scripts containing further configurations for the \ac{Raspi}s are desirable
to be distributed as part of the common image. Therefore, here we present
the steps to setup basic properties of the \ac{Raspi}s and distributing
configuration scripts to each of them through the image. For this, we first
indicate how to obtain and put our configuration scripts in the image. Later,
we describe the tasks performed by these configuration scripts. Finally, we
indicate how and in which order the scripts are executed to configure all the
devices. Any testbed administrator might modify or include other
tasks besides the presented in this procedure according to his / her needs
as we will show.

@@

% MAC and Hostname file
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/nodes.csv}@>
<@\textcolor{gray}{
--------------------------------------------------------------------------------}
\Reactivatenumber @>
# Ethernet MAC    Hostname
b8:27:eb:5b:da:20 rasp00
b8:27:eb:7b:c3:91 rasp01
b8:27:eb:54:9c:64 rasp02
b8:27:eb:95:bd:11 rasp03
...
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

\subsubsection{Updating Default Configuration Files and Scripts}
Besides the single script with its configuration file introduced up
to this point of our procedure, it is possible that the testbed
administrator may require to add other scripts to configure
his / her \ac{Raspi}s. We want
to ensure that all the \ac{Raspi} configuration scripts of any testbed
administrators are obtained in a simple way. We automate this task by
including the \texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/update\_rasp\_config}
script in our procedure. The purpose of this script is to make all the
\ac{Raspi}s fetch all the configuration scripts located with the image
during a testbed start up.

In our case as the testbed administrator for presenting the procedure, we want
to fetch all our configuration scripts in
\texttt{\$\{URL\%/*\}/raspi\_config.zip}. The update script
\textit{automatically} downloads all the required configuration files in
\texttt{rasp\_config.zip} file from a remote location. This is the same that
we manually did earlier to get our files but this will be made in an
automated way after booting up the system. This script content is:

@@

The update script lines (3-4) checks for the \ac{URL} and \texttt{.zip} file that
should be downloaded. Lines (7-10) downloads the configuration files to \texttt{/tmp}
folder in the corresponding \ac{Raspi}. It also prints a warning in case of
errors and line (13) unzips the files to \texttt{/home/pi/}, the \ac{Raspi} home
directory. Existing files and directories are simply overwritten.
%while ignoring existing files and / or directories there.

For the above scripts to work in the \ac{Raspi}s, it is required that the
\ac{Raspi}s \ac{MAC} addresses are found in \texttt{nodes.csv}. Also, it
should be noted that for other testbed administrators besides ourselves,
the \ac{URL} for file fetching and the configuration
scripts themselves can be modified to fit their requirements. If
required for a testbed administrator, the \texttt{rasp\_config.zip}
will need to be edited to include all the required configuration files
and scripts. Also, it might be needed to edit the \ac{URL} in the script
\texttt{update\_rasp\_config} to store and fetch from a different location.
Nevertheless, both the \ac{URL} and configuration files presented here can be
used as a starting boilerplate if desired.

\subsubsection{Configuration Scripts Execution Order}
To actually make the \ac{Raspi}s change hostnames and any other considered
configurations, we have to make each \ac{Raspi} call the above scripts when
it starts up. After finishing the setup process, all the unzipped files
presented in Section~\ref{sec:configuration_files_download} should be
locally available at each \ac{Raspi} after getting the root filesystem.
We first need to run the update script before running
any other configuration scripts. To do this after boot up, we include
a call for the update script in the image \texttt{\$\{ROOTDIR\}/etc/rc.local}
file before \texttt{exit 0} using a file editor. To do this, it is required
to edit this file with root permissions. The resulting file should look like the following:

@@

Provided that qemu-arm is enabled, we should now be able to \texttt{chroot}
into our Raspbian lite image. There are a few commands to be
performed before actually changing root into the root partion of the image.
First, to get internet access from within the Raspbian lite image it is needed
to copy the testbed administrator Linux distribution \texttt{resolv.conf} file
into the image root filesystem. To do this, it is required to run the
following:

@@

\texttt{vim} is the improved \texttt{vi} editor for Linux, \texttt{git}
for managing Git repositories and \texttt{screen}~\cite{gnu_screen} for better handling of
long-runnning processes. When writing the image to a memory card, all the
changes that have been made to the image so far, will exist in all
\ac{Raspi}s after fetching it.

### 2016-09-14 introduction.tex
first round of corrections

The upcoming 5G technology is targeting the controlling and steering of the \ac{IoT} in real-time on global scale. This will break new ground for new markets such as driverless vehicles, manufacturing, humanoid robots, smart grids, and many more. With 5G, the number of wireless devices will increase by five times up to 50 billion devices \cite{cisco2011forecast}. It is generally believed, that those devices will not be connected in the same manner as current devices are connected today. Centralized system will collapse in terms of capacity, while distributed systems appear as an alternative. Therefore the communication architecture in future systems will be dominated by mesh technologies. Mesh technology has been known for sensor and ad hoc networks or mobile cloud scenarios, but the technical requirements on 5G mesh-based communication systems are dramatically increasing. Future mesh networks need to support high data rate, low latency, security, network availability and heterogeneous devices to ensure high \ac{QoE} for the final user. In state of the art systems those requirements are traded-off with each other, but in the 5G context we cannot do this anymore.

@@

In our prior works, the C++11 Kodo library \cite{kodo2011pedersen} was used as the common building block containing the basic \ac{RLNC} functionalities. However, most of the work was focusing on small mesh networks with a handful of communication nodes, though the expected scenarios are fairly beyond this order of magnitude. Moreover, despite this successful deployment in real systems, many of these protocols and contributions have been implemented in separate testbeds and the experiences are hard to reproduce. Furthermore, deploying a large-scale and configurable testbed for networking and storage can be challenging, not only due to the inherent costs of the hardware, but due to maintenance challenges and ability to replicate results consistently. The latter requires not only the devices to run the same \ac{OS}, but also have exactly the same configurations and software packages. Hence, there is a need in evaluating large scale network deployments of low cost devices in a quick, easy-to-deploy, reproducible and maintainable fashion.

The emergence of powerful and inexpensive single-board computers running full versions of a standard \ac{OS} from many possible opens new possibilities in this area. This not only allows to seamlessly develop implementations that are compatible with higher end devices, but also permits the reuse of stable software from widely-used \ac{OS}. For example, the work in \cite{cox2014iridis} provides a detailed description of a \ac{Raspi}~\cite{making_of_pi} testbed ideal for educational applications. Here the authors present computational speed benchmarks, inter-node communication throughput and memory card writing speeds for data storage to assess the testbed performance. However, this work indicates only a basic description of how to set up the required software and also mentions that its maintenance could be time-demading. Moreover, this work does not consider possible network coding applications. Different studies of \ac{IoT} applications consider using the \ac{Raspi} for data processing: In \cite{leccese2014smartcity}, the \ac{Raspi} is the processing unit that coordinates and controls the activity of a public road isle of lamps and reports it to a monitoring center. A use case regarding remote environment surveillance is presented in \cite{leccese2014newimaging}, where the \ac{Raspi} reports air pressure, humidity and temperature of the locations of cultural paintings plus high-resolution images of the paintings themselves. This data is sent to a monitoring center to ensure the preservation of the paintings. Further, authors in \cite{sapes2016finger} utilize the \ac{Raspi} to act as the data server in a finger scanning application that collects the fingerprints. Even though all these applications consider the use of the \ac{Raspi} as a core block, they provide few to no description of their procedures to configure the \ac{Raspi}. Moreover, these applications become cumbersome to maintain as their considered systems could potentially scale when aiming to serve more users. Furthermore, the current way that the data is sent in the considered networks for these \ac{IoT} applications will not be feasible in future 5G systems as mentioned previously.

Given this set of specific needs, in this work we present the design, key step-by-step instructions and mechanisms to setup, configure and maintain an inexpensive testbed using potentially several \ac{Raspi} devices for networking (wireless or wired) and storage applications including \ac{RLNC} functionalities into the testbed through Kodo. Nevertheless, the architecture itself is not bounded to the networking area and can be used for other applications that require replicable results with the \ac{Raspi}. Our work for the testbed procedure is organized as follows: Section~\ref{sec:overview} introduces the testbed system. In Section~\ref{sec:image_setup}, we provide details about the testbed setup, scripts, configuration files and connectivity. In Section~\ref{sec:overlay_fs}, we elaborate on the need and setup for an overlay filesystem for our testbed in order to have both persistent and non-persistent data on it as an optional step. Section~\ref{sec:tools} describes a set of automation and monitoring tools that can be included in the testbed to simplify the execution of routinary and repetitive tasks. Section~\ref{sec:cross_compilation} elaborates on the compilation of the Kodo library for the \ac{Raspi}. Conclusions and future work are reviewed in Section~\ref{sec:conclusions}. Finally, a set of alternative commands, in case the ones presented in this work might not be executed, are discussed in the Appendices.

>>>>

The upcoming 5G technology is targeting the controlling and steering of the \ac{IoT} in real-time on a global scale. This will break new ground for new markets such as driverless vehicles, manufacturing, humanoid robots, and smart grids. The number of wireless devices is expected to increase by five times up to 50 billion devices \cite{cisco2011forecast}. It is generally believed, that those devices will not be connected in the same manner as current devices are connected today. Centralized system will collapse in terms of capacity, while distributed systems appear as an alternative. Therefore the communication architecture in future systems will be dominated by mesh technologies. Mesh technology has been known for sensor and ad hoc networks or mobile cloud scenarios, but the technical requirements on 5G mesh-based communication systems are dramatically increasing. Future mesh networks need to support high data rate, low latency, security, network availability and heterogeneous devices to ensure high \ac{QoE} for the final user. In state of the art systems those requirements are traded-off with each other, but in the 5G context we cannot do this anymore.

@@

In our prior works, the C++11 Kodo library \cite{kodo2011pedersen} was used as the common building block containing the basic \ac{RLNC} functionalities. Most of the work was focusing on small mesh networks with a handful of communication nodes, though the expected scenarios are fairly beyond this order of magnitude. Despite this successful deployment in real systems, many of these protocols and contributions have been implemented in separate testbeds and the experiences are hard to reproduce. Deploying a large-scale and configurable testbed for networking and storage can be challenging, not only due to the inherent costs of the hardware, but due to maintenance challenges and ability to replicate results consistently. The latter requires not only the devices to run the same \ac{OS}, but also have exactly the same configurations and software packages. There is a need in evaluating large scale network deployments of low cost devices in a quick, easy-to-deploy, reproducible and maintainable fashion.

The emergence of powerful and inexpensive single-board computers running full versions of a standard \ac{OS} from many possible opens new possibilities in this area. This not only allows to seamlessly develop implementations that are compatible with higher end devices, but also permits the reuse of stable software from widely-used \ac{OS}. For example, the work in \cite{cox2014iridis} provides a detailed description of a \ac{Raspi}~\cite{making_of_pi} testbed ideal for educational applications. Here the authors present computational speed benchmarks, inter-node communication throughput and memory card writing speeds for data storage to assess the testbed performance. However, this work indicates only a basic description of how to set up the required software and also mentions that its maintenance could be time-demading. Moreover, this work does not consider possible network coding applications. Different studies of \ac{IoT} applications consider using the \ac{Raspi} for data processing: In \cite{leccese2014smartcity}, the \ac{Raspi} is the processing unit that coordinates and controls the activity of a public road isle of lamps and reports it to a monitoring center. A use case regarding remote environment surveillance is presented in \cite{leccese2014newimaging}, where the \ac{Raspi} reports air pressure, humidity and temperature of the locations of cultural paintings plus high-resolution images of the paintings themselves. This data is sent to a monitoring center to ensure the preservation of the paintings. Further, authors in \cite{sapes2016finger} utilize the \ac{Raspi} to act as the data server in a finger scanning application that collects the fingerprints. Even though all these applications consider the use of the \ac{Raspi} as a core block, they provide few to no description of their procedures to configure the \ac{Raspi}. Moreover, these applications become cumbersome to maintain as their considered systems could potentially scale when aiming to serve more users. The current way that the data is sent in the considered networks for these \ac{IoT} applications will not be feasible in future 5G systems as mentioned previously.

Given this set of specific needs, in this work we present the design, key step-by-step instructions and mechanisms to setup, configure and maintain an inexpensive testbed using potentially several \ac{Raspi} devices for networking (wireless or wired) and storage applications including \ac{RLNC} functionalities into the testbed through Kodo. The architecture itself is not bounded to the networking area and can be used for other applications that require replicable results with the \ac{Raspi}. Our work for the testbed procedure is organized as follows: Section~\ref{sec:overview} introduces the testbed system. In Section~\ref{sec:image_setup}, we provide details about the testbed setup, scripts, configuration files and connectivity. In Section~\ref{sec:overlay_fs}, we elaborate on the need and setup for an overlay filesystem for our testbed in order to have both persistent and non-persistent data on it as an optional step. Section~\ref{sec:tools} describes a set of automation and monitoring tools that can be included in the testbed to simplify the execution of routinary and repetitive tasks. Section~\ref{sec:cross_compilation} elaborates on the compilation of the Kodo library for the \ac{Raspi}. Conclusions and future work are reviewed in Section~\ref{sec:conclusions}. Finally, a set of alternative commands, in case the ones presented in this work might not be executed, are discussed in the Appendices.

### 2016-09-14 overlay_fs.tex
first round of corrections

\label{sec:overlay_fs}
In principle, our procedure modifies the image file only once in
the testbed administrator \ac{PC} when its setup is made. Also, keeping
this image in the \ac{Raspi}s provides the same initial system for all
the devices. If we do not make any further modifications during
the image setup, any files created after the initial boot of a \ac{Raspi}
will remain in the memory card. This is cumbersome to maintain since
the size of the memory card is relatively small (8 GB) and there
might be various users utilizing the testbed. Also, different testbed
users could be interested in running their experiments in a fresh
rebooted system with the original customized image. Nevertheless, we
emphasize that this step is not necessary if the tesbed administrator
wants to consider only persistent storage for its devices. A use case
for this scenario could be a single user for the testbed or when a testbed
administrator only wants to setup a few \ac{Raspi}s.

If both persistent and non-persistent storage are required for
the \ac{Raspi}s, here we present the steps to setup an overlay filesystem.
This type of filesystem enables an \textit{upper} filesystem to overlay
into a \textit{lower} filesystem. Whenever a file is requested, the upper
filesystem will forward the request to the lower filesystem in case it
does not have it itself. If the upper filesystem has the requested file,
it will simply return the file. This idea can be used in our setup to mount
the root filesystem (i.e. Raspbian Lite) in the \ac{Raspi}s during startup
as read-only filesystem. Thus, on the one hand, the image configuration files
will remain after a reboot but the local data in these directories will be
erased after a hypothetic reboot. On the other hand, to enable the
possibility of persistent changes, we overlay the upper filesystem that is
mounted in the \ac{Raspi} \ac{RAM}, i.e. \texttt{/tmp} as read / writable
on top of the lower root filesystem. Thus, reading a file may return a
file from the lower filesytem, but if it is stored, it will be saved in
the upper filesystem. Accessing this file again will return
the stored file from the upper layer. After a reboot, all the stored files
in the upper filesystem will be retrieved, but the ones in the lower
filesystem that are not part of the original image will be removed.

@@

Assuming that we are still in the \texttt{chroot} environment of the
Raspbian Lite root filesystem for installing packages, we can setup the
overlay filesystem at this point of the procedure. There already exists
implementations overlaying the root filesystem. We use an implementation
available at the Git repository in \cite{overlayroot}. Since we have
installed \texttt{git} in a previous step, we clone the repository. The
command block below stores it in \texttt{/tmp} which is really mounted in
\ac{RAM}. Thus, all the files stored here will disappear when the system is
rebooted.

@@

Again, if for some reason, the \texttt{-P} option is not available for
\texttt{losetup}, we provide an alternative command line in
Section~\ref{sec:alternative_losetup}. Finally, if the previous filesystem
formatting was successful, the filesystem is now available for use. Thus,
we need to inform the Raspbian \ac{OS} to mount the home partition
that we have just created.
%We will notify the \ac{OS} while booting during the startup of the
%\ac{Raspi}s.
This can be done by adding
an entry in \texttt{fstab} as follows:

@@

Again, if the \texttt{----recursive} option is not available, the
procedure in Section~\ref{sec:umount} of the Appendices could be followed.
If the steps are successfully executed up to this point, the customized image
is available in the \texttt{\$\{IMAGE\}.img} file and ready to be deployed
into the \ac{Raspi}s. In the following section, we indicate how to proceed
with the writing of the image into various memory cards.

>>>>

\label{sec:overlay_fs}
In principle, our procedure modifies the image file only once in
the testbed administrator \ac{PC} when its setup is made. Also, keeping
this image in the \ac{Raspi}s provides the same initial system for all
the devices. If we do not make any further modifications during
the image setup, any files created after the initial boot of a \ac{Raspi}
will remain in the memory card. This is cumbersome to maintain since
the size of the memory card is relatively small (8 GB) and there
might be various users utilizing the testbed. Also, different testbed
users could be interested in running their experiments in a fresh
rebooted system with the original customized image. We
emphasize that this step is not necessary if the tesbed administrator
wants to consider only persistent storage for its devices. A use case
for this scenario could be a single user for the testbed or when a testbed
administrator only wants to setup a few \ac{Raspi}s.

If both persistent and non-persistent storage are required for
the \ac{Raspi}s, here we present the steps to setup an overlay filesystem.
This type of filesystem enables an \textit{upper} filesystem to overlay
into a \textit{lower} filesystem. Whenever a file is requested, the upper
filesystem will forward the request to the lower filesystem in case it
does not have it itself. If the upper filesystem has the requested file,
it will simply return the file. This idea can be used in our setup to mount
the root filesystem (i.e. Raspbian Lite) in the \ac{Raspi}s during startup
as read-only filesystem. On the one hand, the image configuration files
will remain after a reboot but the local data in these directories will be
erased after a reboot. To enable the
possibility of persistent changes, we overlay the upper filesystem that is
mounted in the \ac{Raspi} \ac{RAM}, i.e. \texttt{/tmp} as rewritable
on top of the lower root filesystem. Reading a file may return a
file from the lower filesytem, but if it is stored, it will be saved in
the upper filesystem. Accessing this file again will return
the stored file from the upper layer. After a reboot, all the stored files
in the upper filesystem will be retrieved, but the ones in the lower
filesystem that are not part of the original image will be removed.

@@

Assuming that we are still in the \texttt{chroot} environment of the
Raspbian Lite root filesystem for installing packages, we can setup the
overlay filesystem at this point of the procedure. There already exists
implementations overlaying the root filesystem. We use an implementation
available at the Git repository in \cite{overlayroot}. Since we have
installed \texttt{git} in a previous step, we clone the repository. The
command block below stores it in \texttt{/tmp} which is really mounted in
\ac{RAM}. All the files stored here will disappear when the system is
rebooted.

@@

If the \texttt{-P} option is not available for
\texttt{losetup}, we provide an alternative command line in
Section~\ref{sec:alternative_losetup}. Finally, if the previous filesystem
formatting was successful, the filesystem is now available for use.
We need to inform the Raspbian \ac{OS} to mount the home partition
that we have just created.
%We will notify the \ac{OS} while booting during the startup of the
%\ac{Raspi}s.
This can be done by adding
an entry in \texttt{fstab} as follows:

@@

If the \texttt{----recursive} option is not available, then follow the
procedure in Section~\ref{sec:umount} of the Appendices.
If the steps are successfully executed up to this point, the customized image
is available in the \texttt{\$\{IMAGE\}.img} file and ready to be deployed
into the \ac{Raspi}s. In the following section, we indicate how to proceed
with the writing of the image into various memory cards.
